\documentclass{article}
% This is a LaTeX file.  It is a text file that is compiled
% by a program called LaTeX into a pretty PDF file.  
% If you're viewing this file on CoCalc, you'll see that PDF 
% in the window to the right.
%
% The LaTeX macro language is complicated, so we have inserted
% lots of documenting comments into the file.  Comments start
% with `%' and continue to the end of the line.  In CoCalc's
% window, they are colored brownish-red.
%
% Comments prefixed with `Student:' are relevant to students.
% Skip anything else you don't understand, or ask me.
%
% set font encoding for PDFLaTeX or XeLaTeX
\usepackage{ifxetex}
\ifxetex
  \usepackage{fontspec}
\else
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{lmodern}
\fi

% Student: These lines describe some document metadata.
\title{Problem Set 1}
\author{%
% Student: change the next line to your name!
    Name
\\  MATH-UA 120 Discrete Mathematics
}
\date{due September 15, 2023}


\usepackage[headings=runin-fixed-nr]{exsheets}
% These make enumerates within questions start at the second ("(a)") level, rather than the first ("1.") level.
\makeatletter
    \newcommand{\stepenumdepth}{\advance\@enumdepth\@ne}
\makeatother
\SetupExSheets{
    question/pre-body-hook=\stepenumdepth,
    solution/pre-body-hook=\stepenumdepth,
}
\DeclareInstance{exsheets-heading}{runin-nn-np}{default}{
    runin = true,
    title-post-code = .\space,
    join = {
        main[r,vc]title[l,vc](0pt,0pt);
    }
}
\newif\ifshowsolutions
% Student: replace `false' with `true' to typeset your solutions.
% Otherwise they are ignored!
\showsolutionstrue
\ifshowsolutions
    \SetupExSheets{
        question/pre-hook=\itshape,
        solution/headings=runin-nn-np,
        solution/print=true,
        solution/name=Answer
    }%
    \makeatletter%
    \pretocmd{\@title}{Answers to }%
    \makeatother%
\else
    \SetupExSheets{solution/print=false}
\fi

% Bug workaround: http://tex.stackexchange.com/a/146536/1402
%\newenvironment{exercise}{}{}
\RenewQuSolPair{question}{solution}
%\let\answer\solution
%\let\endanswer\endsolution
\usepackage{manfnt}
\newcommand{\danger}{\marginpar[\hfill\dbend]{\dbend\hfill}}

\usepackage{amsmath, amsthm}
\usepackage{amsfonts}
\usepackage{enumerate}
\usepackage{siunitx}
\DeclareSIUnit\pound{lb}
\usepackage{hyperref}
\newtheorem*{theorem}{Theorem}
\newtheorem*{claim}{Claim}
\theoremstyle{definition}
\newtheorem*{definition}{Definition}
% We are creating a command "\xor".
\newcommand{\xor}{\underline{\lor}}
% This is the beginning of the part of the file that describes
% the text of the document.
% That's why it says `\begin{document}' below. :-)
\begin{document}
\maketitle



These are to be written up and turned in to Gradescope.\\



\ifshowsolutions
    \SetupExSheets{solution/print=true}
\else
    \danger
 \underline{ \LaTeX  Instructions:}  You can view the source (\texttt{.tex}) file to get some more examples of \LaTeX{} code.  I have commented in the source file in places where new \LaTeX{} constructions are used.
  
  Remember to change \verb|\showsolutionsfalse| to \verb|\showsolutionstrue|
    in the document's preamble 
    (between \verb|\documentclass{article}| and \verb|\begin{document}|)
\fi

\section*{Assigned Problems}


\begin{question}
    % Notice the use of the enumerate environment
    % to make a numbered list.  Each item is marked
    % by the \item command.
    %
    % Also \emph = emphasize, usually in italics.
    Let the following statements be given. 
       \begin{definition}
          A triangle is \emph{scalene} if all of its sides have different lengths.
       \end{definition}
       \begin{theorem}
          A triangle is scalene if it is a right triangle that is not isosceles.
       \end{theorem}
    Suppose $\Delta ABC$ is a scalene triangle. 
    Which of the following conclusions are valid based only on the information given above? 
    Why or why not?
    \begin{enumerate}
        \item All of the sides of $\Delta ABC$ have different lengths.
        \item $\Delta ABC$ is a right triangle that is not isosceles.
    \end{enumerate}
\end{question}
% Student: put your answer between the next two lines.
\begin{solution}
    \begin{enumerate}
        \item This is valid based on the definition. Note that definitions are "if and only if" statements.
        \item This is not valid based on the theorem, since we are given the conclusion, 
            we cannot say anything about the hypothesis.
    \end{enumerate}
    
{\color{red} Rubric:
\begin{itemize}
\item 5P per part:
\item 2P for answer
\item 3P for reasoning
\end{itemize}}
\end{solution}


\begin{question}
   Without changing their meanings, convert each of the following sentences into a sentence of the form ``For all ... $x$, if $x$ ... , then .'', regardless of whether the statement is true or false.
    \begin{enumerate}
        \item Every prime greater than 2 is odd.
        \item Three consecutive odd integers greater than 3 cannot all be prime.
        \item An integer is divisible by 8 only if it is divisible by 4.
        \item You fail only if you stop writing.
        \item People will generally accept facts as truth only if the facts agree with what they already believe.
    \end{enumerate}
\end{question}
% Student: put your answer between the next two lines.
\begin{solution}
    \begin{enumerate}
        \item For all integers $x$, if $x>2$ and prime, then $x$ is odd. 
        \item For all integers $x$, if $x$, $x+2$, and $x+4$ are all greater than 3, then they cannot all be prime. 
        \item For all integers $x$, if $x$ is divisible by 8, then $x$ is divisible by 4.
        \item For any person $x$, if $x$ stops writing, then $x$ fails.
        \item For any person $x$, if facts agree with what $x$ already believes, then $x$ will generally accept the facts as truth.
    \end{enumerate}

{\color{red} Rubric:
\begin{itemize}
\item 2P for statement per part
\item Try your best to give partial credit for rewritten statement.
\end{itemize}}
\end{solution}


\begin{question}
   The following claim and its proof is poorly written. They are both missing some crucial information. 
   Please revise both the claim and proof so that any student in this course will understand it. Explain your reasoning.
      \begin{claim}
       If $x^2\neq 0$, then $x^2>0$.
      \end{claim}
      \begin{proof}
       If $x>0$, then $x^2=xx>0$. If $x<0$, then $-x>0$, so $(-x)(-x)>0$, i.e., $x^2>0$.
      \end{proof}
\end{question}
% Student: put your answer between the next two lines.
\begin{solution}
    The claim has no information about the quantity $x$. A clearer statement would be:
        \begin{claim}
           For all real numbers $x$, if $x^2\neq 0$, then $x^2>0$.
        \end{claim}
    In the proof, the assumption $x^2\neq 0$ is missing. 
    From there, it looks like the original proof is broken into cases as a result of the hypothesis. 
    Here is a clearer version of the proof:
        \begin{proof}
           Let $x$ be a real number such that $x^2\neq 0$. Then $x\neq 0$, and we have two cases:
            \begin{enumerate}[i.]
                 \item if $x>0$, then $x\cdot x=x^2>0$.
                 \item if $x<0$, then $-x>0$, so $(-x)(-x)>0$, that is, $x^2>0$.
            \end{enumerate}
           Therefore, in both cases, $x^2>0$.
        \end{proof}

{\color{red} Rubric:
\begin{itemize}
\item Claim: 2P for fix + 2P for reasoning
\item Proof: 4P for fix + 2P for reasoning
\end{itemize}}
\end{solution}


\begin{question}
    Consider the following definition of the ``$\triangleleft$'' symbol.
	\begin{definition}
	 Let $x$ and $y$ be integers. Write $x\triangleleft y$ if $3x+5y=7k$ for some integer $k$.
	\end{definition}
        \begin{enumerate}
           \item Show that $1\triangleleft 5$, $3\triangleleft 1$, and $0\triangleleft 7$.
           \item Prove that if $a\triangleleft b$ and $c\triangleleft d$, then $(a+c) \triangleleft (b+d)$.
        \end{enumerate}
\end{question}
% Student: put your answer between the next two lines.
\begin{solution}
    \begin{enumerate}
      \item We want to find an integer $k$ that satisfies $3x+5y=7k$.
	\begin{itemize}
	\item For $1\triangleleft 5$, we need $k=4$ such that $3(1)+5(5)=7(4)$.
	\item For $3\triangleleft 1$, we need $k=2$ such that $3(3)+5(1)=7(2)$.
	\item For $0\triangleleft 7$, we need $k=5$ such that $3(0)+5(7)=7(5)$.
	\end{itemize}
      \item Suppose that $a\triangleleft b$ and $c\triangleleft d$. Then there exists integers $k_1$ and $k_2$ such that $3a+5b=7k_1$ and $3c+5d=7k_2$.
	    \begin{eqnarray*}
	    (3a+5b) + (3c+5d) &=& 7k_1 + 7k_2\\
	    3(a+c) + 5(b + d) &=& 7(k_1+k_2)
	    \end{eqnarray*}
	Let $k=k_1+k_2$. Since there exists an integer $k$ such that 
	$$3(a+c) + 5(b + d)=7k,$$ $(a+c) \triangleleft (b+d)$.
    \end{enumerate}
    
{\color{red} Rubric:
\begin{itemize}
\item 1P for part a
\item Part b: follow RVF 9P rubric
\end{itemize}}
\end{solution}

\begin{question}
    Show that an integer $n$ is odd if and only if $2n+2$ is divisible by 4.
\end{question}
% Student: put your answer between the next two lines.
\begin{solution}
    This statement can be written as 
	\begin{center}
		$n \in \mathbb{Z}$ is odd $\iff 4 ~|~ (2n+2)$.
	\end{center}
    ($\Rightarrow$) Assume first that $n$ is odd. Then we can write $n = 2k+1$ for some integer $k$. Then
	\[
	2 n + 2 = 2 (2k + 1) + 2 = 4 k + 2 + 2 = 4 k + 4 = 4 (k+1).
	\]
	Since $k+1 \in \mathbb{Z}$, it means by definition that $2 n + 2$ is divisible by 4.
	
    ($\Leftarrow$)  Assume now that $2n + 2$ is divisible by 4. 
    Then we can write $2n + 2 = 4k$ for some $k \in \mathbb{Z}$. Simplifying gives $n + 1 = 2k$, and thus
	\[
	n = 2 k - 1 = 2 (k - 1) + 1.
	\]
    Since $k - 1 \in \mathbb{Z}$, this means by definition that $n$ is odd.
    
{\color{red} Rubric:
\begin{itemize}
\item Follow RVF 9P rubric 
\item 1P for proper use of \LaTeX to indicate the two directions, or clearly stating in words.
\end{itemize}}
\end{solution}


\end{document}
